---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tvrestoreR

<!-- badges: start -->
<!-- badges: end -->

`tvRestoreR` is an R package for image restoration using total variation (TV) regularization.  
It supports inpainting (filling in missing or masked pixels) and denoising (removing noise from observed pixels) through a single variational framework, and ROF-inpainting (equivalent to inpainting and denoising concurrently). 
Using a primal-dual (Chambolle–Pock) solver, the package takes a user-provided image and mask, applies TV regularization with a tuning parameter (λ), and returns a restored image that respects the observed data while enforcing smoothness and preserving edges.


## Installation

You can install the development version of tvrestoreR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jaylee0717-dev/tvrestoreR")
```

## Mathematical Formulation
The package solves the following minimization problem:$$ \min_{u} \int_{\Omega} |\nabla u| + \frac{\lambda}{2} \int_{\Omega \cap D} (u - f)^2 dx $$

Where:$u$ is the restored image.
$\int |\nabla u|$ is the Total Variation term (enforces smoothness/sparsity of gradients).
$f$ is the observed noisy image.
$D$ is the domain of observed pixels (the mask).
$\lambda$ is the regularization parameter.

there are two implemented solver methods for the problem: primal-dual by Chambolle-Pock, and split bregman. 

## Example
The following example outlines how to generate synthetic data (including the base "ground-truth" image, corrupted image, and the mask), then run TV restoration, and then visualize. 

```{r}
library(tvrestoreR)

## Generate Synthetic Data 
# --------------------------
# Creates a "wedge" pattern with gaussian noise with SD 0.1 and 40% missing pixels
sim_data <- generate_example_images(
  pattern = "wedge", 
  noise_sigma = 0.1, 
  missing_fraction = 0.1,
  mask_type = "disk",
  seed = 123
)


## Run TV Restoration
# ---------------------
result <- tv_restore(
  img = sim_data$original,
  mask = sim_data$mask, 
  lambda = 1,
  task = "ROF_inpainting"
) 

## Visualize
# ------------
# Red pixels indicate missing data (NA)
plot_images(
  imgs = list(sim_data$original, sim_data$corrupted, result),
  titles = c("Ground Truth", "Corrupted Input", "TV Restored"),
  na_color = "red"
)
```



