---
title: "Introduction to TV Restoration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to TV Restoration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction 
The tvrestoreR package implements Total Variation (TV) regularization. This vignette demonstrates three core capabilities: Denoising, Inpainting, and ROF-inpainting (performs both denoising and inpainting at the same time). 

```{r}
library(tvrestoreR)
```

## 1. Image Denoising
Denoising attempts to remove Gaussian noise while preserving sharp edges. We will use the generate_example_images() helper to create a "wedge" pattern with noise but no missing pixels.

```{r}
# Generate data: Noise only, no missing pixels
data_denoise <- generate_example_images(
  pattern = "gradient", 
  noise_sigma = 0.2, 
  missing_fraction = 0.0,
  seed = 42
)

# Run Restoration
denoised <- tv_restore(
  img = data_denoise$corrupted, 
  mask = data_denoise$mask, 
  lambda = 1
)

# Visualize
plot_images(
  list(data_denoise$original, data_denoise$corrupted, denoised),
  titles = c("Truth", "Noisy", "Restored")
)
```

## 2. Image Inpainting
Inpainting recovers missing pixels. We will simulate damage by masking out a large "disk" in the center of the image.
```{r}
# Generate data: Wedge pattern with a missing hole
data_inpaint <- generate_example_images(
  pattern = "wedge", 
  noise_sigma = 0.1,        
  missing_fraction = 0.3,    
  mask_type = "disk",
  seed = 42
)

# Run Restoration
inpainted <- tv_restore(
  img = data_inpaint$corrupted, 
  mask = data_inpaint$mask, 
  lambda = 5,
  task = "inpainting"
)

# Visualize
# The red area in the middle represents the missing data
plot_images(
  list(data_inpaint$original, data_inpaint$corrupted, inpainted),
  titles = c("Truth", "Masked Input", "Inpainted"),
  na_color = "red"
)
```


## 3. ROF-inpainting
ROF-inpainting performs both denoising and inpainting. To demonstrate its denoising capability, we use the same image as in part 2. Typically, the optimal lambda is less for ROF-inpainting than the other tasks. 
```{r}
# Run Restoration
ROF_inpainted <- tv_restore(
  img = data_inpaint$corrupted, 
  mask = data_inpaint$mask, 
  task = "ROF_inpainting",
  lambda = 1
)

# Visualize
# The red area in the middle represents the missing data
plot_images(
  list(data_inpaint$original, data_inpaint$corrupted, ROF_inpainted),
  titles = c("Truth", "Masked Input", "ROF-Inpainted"),
  na_color = "red"
)
```

## 4. The Effect of Lambda
The regularization parameter $\lambda$ controls the trade-off between the data fidelity (trusting the input) and the regularization term (smoothness).
High $\lambda$: Strong regularization. The result is very smooth.
Low $\lambda$: Weak regularization. The result stays very close to the noisy input.

```{r}
# Run denoising example with diff lambdas
res_smooth <- tv_restore(data_denoise$corrupted, data_denoise$mask, lambda = 0.01)
res_noisy  <- tv_restore(data_denoise$corrupted, data_denoise$mask, lambda = 10)

plot_images(
  list(res_smooth, res_noisy),
  titles = c("Lambda = 0.01 (Under-smoothed)", "Lambda = 10 (Over-smoothed)"),
  cols = 2
)
```

## 3. Finding the Optimal Lambda
The regularization parameter $\lambda$ controls the trade-off between smoothness and data fidelity. The find_lambda function automates the gridsearch for the best $\lambda$ by minimizing the Mean Squared Error (MSE) against a known ground truth (or a reference image).

```{r}
# Use the noisy wedge data from Section 2
# Define a grid of lambdas to test
l_grid <- c(0.01, 0.1, 0.5, 1, 5, 10, 50)

tuning_results <- find_lambda(
  img = data_inpaint$corrupted,
  mask = data_inpaint$mask,
  ground_truth = data_inpaint$original,
  lambda_grid = l_grid,
  task = "ROF_inpainting",
  method = "primal_dual",
  verbose = FALSE
)

# Plot the Error Curve 
plot(tuning_results$lambda_grid, tuning_results$errors, type='b', 
     col="blue", pch=19, xlab="Lambda", ylab="MSE", main="Parameter Tuning")
grid()
points(tuning_results$best_lambda, min(tuning_results$errors), col="red", pch=19, cex=1.5)
text(tuning_results$best_lambda, min(tuning_results$errors), 
     labels=paste("Best:", tuning_results$best_lambda), pos=3, col="red")
```

## 5. Batch Processing (List of Images)
tv_restore can handle a list of matrices, automatically pre-computing the heavy operators ($D$ and $F$) only once to save time. This is ideal for processing video frames or medical slices.
```{r}
# Create a list of 3 identical noisy frames for demonstration
data_denoise1 <- generate_example_images(
  pattern = "gradient", 
  noise_sigma = 0.2, 
  missing_fraction = 0.0,
  seed = 42
)
data_denoise2 <- generate_example_images(
  pattern = "wedge", 
  noise_sigma = 0.2, 
  missing_fraction = 0.0,
  seed = 42
)
data_denoise3 <- generate_example_images(
  pattern = "wedge60", 
  noise_sigma = 0.2, 
  missing_fraction = 0.0,
  seed = 42
)
batch_input <- list(
  data_denoise1$corrupted, 
  data_denoise2$corrupted, 
  data_denoise3$corrupted
)

# Run batch restoration
# The function detects the list and iterates efficiently
batch_output <- tv_restore(
  img = batch_input,
  mask = data_denoise$mask,
  lambda = 1,
  task = "denoising",
  verbose = FALSE 
)

# Check results
class(batch_output) # Should be "list"
length(batch_output) # Should be 3
```

## 6. Alternative Solver: Split-Bregman
By default, tv_restore uses the Primal-Dual (Chambolle-Pock) algorithm. However, you can also use the Split-Bregman method.

Primal-Dual: Often faster for simple TV problems, very robust.

Split-Bregman: Can be more accurate for specific types of constraints or harder problems.


```{r}
# Compare the two solvers on the inpainting task
res_pd <- tv_restore(
  img = data_inpaint$corrupted, 
  mask = data_inpaint$mask, 
  lambda = 10, 
  method = "primal_dual",
  task = "inpainting",
  verbose = FALSE
)

res_sb <- tv_restore(
  img = data_inpaint$corrupted, 
  mask = data_inpaint$mask, 
  lambda = 10, 
  method = "split_bregman",
  task = "inpainting",
  verbose = FALSE
)

# Visualize the difference
# We plot the absolute difference between the two solutions
diff_img <- abs(res_pd - res_sb)

plot_images(
  list(res_pd, res_sb, diff_img),
  titles = c("Primal-Dual", "Split-Bregman", "Difference"),
  cols = 3
)
```

